---
title: "Harmonizing Cell-type Labels"
author: "<h5>Author: <i>Brian M. Schilder</i></h5>"
date: "<h5>Most recent update: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h5>"
output:
    BiocStyle::html_document:
vignette: >
    %\VignetteIndexEntry{harmonise_celltypes}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 5)
```

# Introduction

Cell-type labels from different datasets are often inconsistent. Authors
use different naming conventions, levels of granularity, and
terminology. For example:

-   "Purkinje neurons" vs "Purkinje cells" vs "PCs"
-   "Astrocytes" vs "Astrocyte-like cells" vs "GFAP+ glia"
-   "Inhibitory neurons" vs "GABAergic neurons" vs "Interneurons"

`scNLP` provides NLP-based approaches to identify semantically similar
labels and harmonize cell-type annotations across datasets without
requiring manual curation.

```{r load-packages, message=FALSE}
library(scNLP)
```

# Example Data

We'll use the included pseudo-bulk dataset containing samples from
multiple studies with heterogeneous cell-type labels:

```{r load-data}
data("pseudo_seurat")

# View the diversity of labels and sources
label_summary <- table(
    Dataset = pseudo_seurat$batch,
    Species = pseudo_seurat$species
)
print(label_summary)
```

```{r unique-labels}
# Count unique cell-type labels
length(unique(pseudo_seurat$celltype))

# Sample of label heterogeneity
head(unique(pseudo_seurat$celltype), 20)
```

# Strategy 1: TF-IDF Clustering

TF-IDF analysis identifies enriched terms within each cluster. Cells
with similar enriched terms likely represent the same cell type,
regardless of the exact label used.

## Compute TF-IDF

```{r tfidf-analysis}
# Run TF-IDF on the existing clusters
result <- run_tfidf(
    obj = pseudo_seurat,
    reduction = "umap",
    cluster_var = "cluster",
    label_var = "celltype"
)
```

## Visualize Enriched Terms

The scatter plot shows cells colored by cluster, with enriched terms
labeled at cluster centers:

```{r tfidf-plot, fig.height=6}
res <- plot_tfidf(
    obj = pseudo_seurat,
    label_var = "celltype",
    cluster_var = "cluster",
    show_plot = TRUE
)
```

## Examine TF-IDF Results

The TF-IDF results reveal which terms are most distinctive for each
cluster:

```{r tfidf-results}
# Get the full TF-IDF results
tfidf_df <- res$tfidf_df

# Top terms per cluster
top_terms <- tfidf_df |>
    dplyr::group_by(cluster) |>
    dplyr::slice_max(order_by = tf_idf, n = 3) |>
    dplyr::select(cluster, word, tf_idf)

print(top_terms, n = 30)
```

Clusters sharing similar enriched terms (e.g., "neuron", "astrocyte",
"oligodendrocyte") likely contain the same cell types from different
datasets.

# Strategy 2: Neighbor-based Harmonization

The `search_neighbors()` function finds cells with similar expression
profiles, allowing you to identify which labels from different datasets
correspond to the same cell types.

## Find Similar Cells

Search for cells similar to those labeled "purkinje":

```{r neighbor-search}
purkinje_neighbors <- search_neighbors(
    seurat = pseudo_seurat,
    var1_search = "purkinje",
    max_neighbors = 10
)

head(purkinje_neighbors)
```

## Cross-dataset Mapping

Examine which labels from other datasets map to Purkinje cells:

```{r neighbor-mapping}
# Get unique neighbor labels
neighbor_labels <- unique(purkinje_neighbors$Var2)
cat("Labels similar to 'purkinje':\n")
print(neighbor_labels)
```

## Search Multiple Cell Types

You can search for multiple cell types to build a harmonization map:

```{r multi-search}
# Search for astrocyte-like cells
astro_neighbors <- search_neighbors(
    seurat = pseudo_seurat,
    var1_search = "astro",
    max_neighbors = 5
)

cat("\nLabels similar to 'astro':\n")
print(unique(astro_neighbors$Var2))
```

# Strategy 3: Combined Approach

Combine TF-IDF with neighbor analysis for robust harmonization.

## Identify Clusters by Enriched Terms

```{r combined-approach}
# Find clusters enriched for neuronal terms
neuronal_terms <- c("neuron", "neuronal", "glutamatergic", "gabaergic")

neuronal_clusters <- tfidf_df |>
    dplyr::filter(word %in% neuronal_terms) |>
    dplyr::group_by(cluster) |>
    dplyr::summarise(
        neuronal_score = sum(tf_idf),
        top_term = word[which.max(tf_idf)]
    ) |>
    dplyr::arrange(dplyr::desc(neuronal_score))

print(neuronal_clusters)
```

## Validate with Neighbor Search

Use neighbor search to validate that cells in these clusters are
transcriptomically similar:

```{r validate-clusters}
# Get cells from a neuronal cluster
if (nrow(neuronal_clusters) > 0) {
    top_cluster <- neuronal_clusters$cluster[1]
    is_top <- pseudo_seurat$cluster == top_cluster
    cluster_cells <- pseudo_seurat$celltype[is_top]
    cat("Labels in top neuronal cluster:\n")
    print(table(cluster_cells))
}
```

# Visualizing Cross-dataset Agreement

## Color by Dataset

Visualize how cells from different datasets intermix:

```{r plot-by-dataset, fig.height=6}
res_batch <- plot_tfidf(
    obj = pseudo_seurat,
    label_var = "celltype",
    cluster_var = "cluster",
    color_var = "batch",
    show_plot = TRUE
)
```

Clusters with cells from multiple datasets that share enriched terms
indicate successful cross-dataset harmonization.

## Color by Species

For cross-species comparisons:

```{r plot-by-species, fig.height=6}
res_species <- plot_tfidf(
    obj = pseudo_seurat,
    label_var = "celltype",
    cluster_var = "cluster",
    color_var = "species",
    show_plot = TRUE
)
```

# Word Cloud Summary

Generate a word cloud to summarize the most distinctive terms across all
clusters:

```{r wordcloud, fig.height=6}
wc <- wordcloud_tfidf(
    obj = pseudo_seurat,
    label_var = "celltype",
    cluster_var = "cluster",
    terms_per_cluster = 5
)
```

# Practical Recommendations

1.  **Start with TF-IDF**: Identify enriched terms to understand the
    semantic content of each cluster.

2.  **Use neighbor search for validation**: Confirm that cells with
    similar labels are transcriptomically similar.

3.  **Examine cross-dataset mixing**: Well-harmonized cell types should
    have cells from multiple datasets in the same cluster.

4.  **Iterate**: Refine cluster assignments and re-run TF-IDF to improve
    harmonization.

5.  **Document mappings**: Create a mapping table of original labels to
    harmonized labels for reproducibility.

# Session Info

<details>

```{r session-info}
utils::sessionInfo()
```

</details>
