---
title: "scNLP: Introduction"
author: "<h5>Author: <i>Brian M. Schilder</i></h5>"
date: "<h5>Most recent update: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h5>"
output:
    BiocStyle::html_document:
vignette: >
    %\VignetteIndexEntry{scNLP}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 5)
```

# Introduction

`scNLP` is an R package for applying Natural Language Processing (NLP)
techniques to single-cell omics data. The primary use case is
harmonizing non-standardized cell-type labels across datasets.

When integrating single-cell datasets from multiple sources, cell-type
annotations are often inconsistent. Different labs may use different
naming conventions for the same cell types (e.g., "Purkinje neurons" vs
"Purkinje cells" vs "PCs"). `scNLP` provides tools to identify
semantically similar labels and find enriched terms that characterize
each cluster.

## Key Features

-   **TF-IDF Analysis**: Identify enriched terms in cell-type labels
    using Term Frequency-Inverse Document Frequency
-
-   **Visualization**: Generate scatter plots and word clouds showing
    enriched terms per cluster
-   **Neighbor Search**: Find cells with similar labels using graph-based
    methods
-   **Seurat Integration**: Seamless integration with the Seurat
    ecosystem

## Installation

```{r install, eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("scNLP")
```

# Quick Start

```{r load-package}
library(scNLP)
```
## Example Data

`scNLP` includes a pseudo-bulk Seurat object containing mean expression
per cell-type from 11 different datasets.

```{r load-data}
data("pseudo_seurat")
pseudo_seurat
```

The metadata contains cell-type labels, batch information, and cluster
assignments:

```{r show-metadata}
head(pseudo_seurat[[]][, c("celltype", "batch", "cluster")])
```

# Core Workflow: TF-IDF Analysis

The core functionality of `scNLP` centers on TF-IDF (Term
Frequency-Inverse Document Frequency) analysis. This NLP technique
identifies words or phrases enriched in one "document" (cluster)
relative to others.

## Understanding TF-IDF

In the context of single-cell data:

-   **Terms**: Words within cell-type labels (e.g., "neuron",
    "astrocyte", "inhibitory")
-   **Documents**: Clusters of cells
-   **Goal**: Find terms that best characterize each cluster

This is analogous to finding marker genes, but for text labels instead
of gene expression.

## Running TF-IDF

The `run_tfidf()` function computes TF-IDF scores and adds results to
the Seurat object metadata:

```{r run-tfidf}
result <- run_tfidf(
    obj = pseudo_seurat,
    reduction = "umap",
    cluster_var = "cluster",
    label_var = "celltype"
)

# View enriched words per cluster
head(result[[]][, c("cluster", "celltype", "enriched_words", "tf_idf")])
```

## Visualizing Results

### Scatter Plot

`plot_tfidf()` creates a dimensional reduction plot with enriched terms
labeled at cluster centers:
```{r plot-tfidf, fig.height=6}
res <- plot_tfidf(
    obj = pseudo_seurat,
    label_var = "celltype",
    cluster_var = "cluster",
    show_plot = TRUE
)
```

The returned object contains:

-   `data`: Processed data used for plotting
-   `tfidf_df`: Full per-cluster TF-IDF results
-   `plot`: The ggplot object

### Word Cloud

For an alternative visualization, `wordcloud_tfidf()` displays enriched
terms as a word cloud:

```{r wordcloud, fig.height=6}
wc_res <- wordcloud_tfidf(
    obj = pseudo_seurat,
    label_var = "celltype",
    cluster_var = "cluster",
    terms_per_cluster = 5
)
```

# Additional Functions

## Preprocessing Pipeline

If starting from raw counts, `seurat_pipeline()` provides a standard
preprocessing workflow:

```{r seurat-pipeline, eval=FALSE}
# Create Seurat object from counts
counts <- Seurat::GetAssayData(pseudo_seurat, layer = "counts")
obj <- SeuratObject::CreateSeuratObject(counts = counts)

# Run standard pipeline
processed <- seurat_pipeline(obj)
```

## Neighbor Search

`search_neighbors()` finds cells with similar labels using the
nearest-neighbor graph:

```{r search-neighbors}
neighbors <- search_neighbors(
    seurat = pseudo_seurat,
    var1_search = "purkinje",
    max_neighbors = 5
)
head(neighbors)
```

## Variable Feature Selection

`FindVariableFeatures_split()` identifies variable features within data
subsets, useful for batch-aware feature selection:

```{r var-features, eval=FALSE}
var_features <- FindVariableFeatures_split(
    seurat = pseudo_seurat,
    split.by = "batch",
    nfeatures = 100,
    nfeatures_max = 50
)
```

# Working with Different Input Types

`scNLP` functions accept multiple input types:
## Seurat Objects

```{r seurat-input, eval=FALSE}
res <- plot_tfidf(
    obj = pseudo_seurat,
    label_var = "celltype",
    cluster_var = "cluster"
)
```

## SingleCellExperiment Objects

```{r sce-input, eval=FALSE}
data("pseudo_sce")
res <- plot_tfidf(
    obj = pseudo_sce,
    label_var = "celltype",
    cluster_var = "cluster"
)
```

## Named Lists

For custom data structures, provide a list with metadata and embeddings:

```{r list-input, eval=FALSE}
data_list <- list(
    metadata = pseudo_seurat[[]],
    embeddings = Seurat::Embeddings(pseudo_seurat, "umap")
)
res <- plot_tfidf(
    obj = data_list,
    label_var = "celltype",
    cluster_var = "cluster"
)
```

# Further Reading

For more detailed examples, see the additional vignettes:

-   `vignette("tf-idf", package = "scNLP")`: Comprehensive TF-IDF
    tutorial
-   `vignette("harmonise_celltypes", package = "scNLP")`: Advanced
    cell-type harmonization workflows

# Session Info

<details>

```{r session-info}
utils::sessionInfo()
```

</details>
