---
title: "tf-idf"
author: "<h5>Author: <i>Brian M. Schilder</i></h5>"
date: "<h5>Most recent update: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h5>"
output:
    BiocStyle::html_document:
vignette: >
    %\VignetteIndexEntry{tf-idf}
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup-options, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 5)
```

# Intro

Term frequency-inverse document frequency (**tf-idf**) is an NLP technique
to identify words or phrases that are enriched in one document relative to
some other larger set of documents.

In our case, our words are within the non-standardized cell labels and our
"documents" are the clusters. The goals is to find words that are enriched
in each cluster relative to all the other clusters. This can be thought of
as an NLP equivalent of finding gene markers for each cluster.


# Examples

```{r load-library, include=TRUE}
library(scNLP)
data("pseudo_seurat")
```

## Preprocessing

If you don't already have a `Seurat` object with reduced dimensions and
cluster assignments, you can generate a new one with the following support
function.

```{r create-seurat}
## Create a fresh Seurat object from raw data
counts <- Seurat::GetAssayData(pseudo_seurat, layer = "counts")
meta.data <- pseudo_seurat[[]]

## Create Seurat object with metadata, then run pipeline
obj <- SeuratObject::CreateSeuratObject(
    counts = counts,
    meta.data = meta.data
)
processed_seurat <- seurat_pipeline(obj = obj)
```


## td-idf annotation

`seurat_tfidf` will run **tf-idf** on each cluster and put the results in
the **enriched_words** and **tf_idf** cols of the `meta.data`.

```{r run-tfidf}
pseudo_seurat_tfidf <- run_tfidf(
    obj = pseudo_seurat,
    reduction = "UMAP",
    cluster_var = "cluster",
    label_var = "celltype"
)
head(pseudo_seurat_tfidf[[]])
```

## td-idf scatter plot

You can also plot the results in reduced dimensional space (e.g. UMAP).
`plot_tfidf()` will produce a list with three items.
- `data`: The processed data used to create the plot.
- `tfidf_df`: The full per-cluster TF-IDF enrichment results.
- `plot`: The `ggplot`.

### `Seurat` input

```{r plot-tfidf}
res <- plot_tfidf(
    obj = pseudo_seurat,
    label_var = "celltype",
    cluster_var = "cluster",
    show_plot = TRUE
)
```

You can color the point by other metadata attributes instead.

```{r plot-by-batch}
res <- plot_tfidf(
    obj = pseudo_seurat,
    label_var = "celltype",
    cluster_var = "cluster",
    color_var = "batch",
    show_plot = TRUE
)
```


### `SingleCellExperiment` input

`plot_tfidf()` can also take in an object of class `SingleCellExperiment`.

```{r sce-input, eval=FALSE}
data("pseudo_sce")

res <- plot_tfidf(
    obj = pseudo_sce,
    label_var = "celltype",
    cluster_var = "cluster",
    show_plot = TRUE
)
```


### `list` input

Lastly, if your data doesn't fit the above example data types, you can
simply supply a named `list` with **metadata** and **embeddings**.

```{r list-input, eval=FALSE}
sce_coldata <- SingleCellExperiment::colData(pseudo_sce)
data_list <- list(
    metadata = sce_coldata,
    embeddings = sce_coldata[, c("UMAP.1", "UMAP.2")]
)

res <- plot_tfidf(
    obj = data_list,
    label_var = "celltype",
    cluster_var = "cluster",
    show_plot = TRUE
)
```

### Interactive mode

You can also create an interactive version of this plot.

```{r interactive-plot, warning=FALSE, eval=FALSE}
res <- plot_tfidf(
    obj = pseudo_seurat_tfidf,
    label_var = "celltype",
    cluster_var = "cluster",
    interact = TRUE,
    show_plot = TRUE,
    species = "species",
    dataset = "dataset",
    enriched_words = "enriched_words",
    tf_idf = "tf_idf"
)
```

# tf-idf wordcloud

You can also show the per-cluster **tf-idf** results as a wordcloud.

```{r wordcloud}
wordcloud_res <- wordcloud_tfidf(
    obj = pseudo_seurat,
    label_var = "celltype",
    cluster_var = "cluster",
    terms_per_cluster = 10
)
print(wordcloud_res$tfidf_df)
```

# Session Info

<details>

```{r Session Info}
utils::sessionInfo()
```

</details>
