---
title: "tf-idf"
author: "<h5>Author: <i>Brian M. Schilder</i></h5>" 
date: "<h5>Most recent update: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h5>"
output:
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
    df_print: paged
vignette: >
    %\VignetteIndexEntry{GWAS_Atlas} 
    %\usepackage[utf8]{inputenc}
    %\VignetteEngine{knitr::rmarkdown} 
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = T, fig.width = 7, fig.height = 5)
```

# Intro 

Term frequency–inverse document frequency (**tf-idf**) is an NLP technique to identify words or phrases that are enriched in one document relative to some other larger set of documents. 

In our case, our words are within the non-standardized cell labels and our "documents" are the clusters. The goals is to find words that are enriched in each cluster relative to all the other clusters. This can be thought of as an NLP equivalent of finding gene markers for each cluster.  


# Examples 

```{r setup, include=T}
library(scNLP) 
data("pseudo_seurat")
```

## Preprocessing 

If you don't already have a `Seurat` object with reduced dimensions and cluster assignments, you can generate a new one with the following support function.

```{r}
## Create some mock raw data 
counts <- Seurat::GetAssayData(pseudo_seurat)
meta.data <- pseudo_seurat@meta.data

processed_seurat <- seurat_pipeline(counts = counts, 
                                    meta.data = meta.data)
```


## td-idf annotation  

`seurat_tfidf` will run **tf-idf** on each cluster and put the results in the **enriched_words** and **tf_idf** cols of the `meta.data`. 

```{r} 
pseudo_seurat_tfidf <- run_tfidf(object = pseudo_seurat,
                                 reduction = "UMAP",
                                 cluster_var = "cluster",
                                 label_var = "celltype") 
head(pseudo_seurat_tfidf@meta.data)
```

## td-idf scatter plot  

You can also plot the results in reduced dimensional space (e.g. UMAP). 
`plot_tfidf()` will produce a list with three items.
- `data`: The processed data used to create the plot. 
- `tfidf_df`: The full per-cluster TF-IDF enrichment results.
- `plot`: The `ggplot`. 

### `Seurat` input

```{r}
res <- plot_tfidf(object = pseudo_seurat, 
                  label_var = "celltype", 
                  cluster_var = "cluster", 
                  show_plot = T)
```

You can color the point by other metadata attributes instead.

```{r}
res <- plot_tfidf(object = pseudo_seurat, 
                  label_var = "celltype", 
                  cluster_var = "cluster", 
                  color_var = "batch",
                  show_plot = T)
```


### `SingleCellExperiment` input 

`plot_tfidf()` can also take in an object of class `SingleCellExperiment`.

```{r not knitting for some reason, eval=F}  
data("pseudo_sce")

res <- plot_tfidf(object = pseudo_sce, 
                  label_var = "celltype", 
                  cluster_var = "cluster", 
                  show_plot = T)
```


### `list` input  

Lastly, if your data doesn't fit the above example data types, you can simply supply a named `list` with **metadata** and **embeddings**. 

```{r} 
data_list <- list(metadata = SingleCellExperiment::colData(pseudo_sce),
                  embeddings = SingleCellExperiment::colData(pseudo_sce)[,c("UMAP.1","UMAP.2")])

res <- plot_tfidf(object = data_list,
                  label_var = "celltype", 
                  cluster_var = "cluster", 
                  show_plot = T)
```

### Interactive mode 

You can also create an interactive version of this plot. 

```{r, warning=F}
res <- plot_tfidf(object = pseudo_seurat_tfidf,
                  label_var = "celltype", 
                  cluster_var = "cluster", 
                  interact = T,
                  show_plot = T,
                  ### Add other metadata vars you want in the hover label like so:
                  species="species",
                  dataset="dataset", 
                  enriched_words="enriched_words",
                  tf_idf="tf_idf") 
```

# tf-idf wordcloud 

You can also show the per-cluster **tf-idk** results as a wordcloud. 

```{r}
wordcloud_res <- wordcloud_tfidf(object=pseudo_seurat,
                                 label_var = "celltype", 
                                 cluster_var = "cluster", 
                                 terms_per_cluster=10)
print(wordcloud_res$tfidf_df)
```

 
# Session Info 

<details> 

```{r Session Info}
utils::sessionInfo()
```

</details>  


