---
title: "hexSticker"
date: "<h4>Updated: <i>`r format( Sys.Date(), '%b-%d-%Y')`</i></h4>"
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{hexSticker}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo=FALSE, include=TRUE}
#### Install deps ####
for(p in c("hexSticker","dplyr","here")){
    if(!require(p,character.only = TRUE)) install.packages(p)     
}
#### Get package metadata deps ####
pkg <- read.dcf(here::here("DESCRIPTION"), fields = "Package")[1]
description <- read.dcf(here::here("DESCRIPTION"), fields = "Description")[1]
```

# Create hex sticker

Hex stickers are a great way to give your R package its very own logo. 
See [here for some examples](http://hexb.in/) from other packages. 

Here are some R packages that you may find helpful for making hex stickers:

## [`hexSticker`](https://github.com/GuangchuangYu/hexSticker)  
Helper functions for creating reproducible hexagon sticker purely in R.

## [`ggimage`](https://github.com/GuangchuangYu/ggimage) 
Supports image files and graphic objects to be visualized in `ggplot2` graphic system.

## [`ggpattern`](https://coolbutuseless.github.io/package/ggpattern/)
Custom `ggplot2` geoms which support filled areas with geometric and image-based patterns.

## [`magick`](https://cran.r-project.org/web/packages/magick/vignettes/intro.html)
Advanced Image-Processing in R.

## [`aRtsy`](https://github.com/koenderks/aRtsy)
R package for making generative art using `ggplot2`.

## [`threed`](https://github.com/coolbutuseless/threed)
Three-Dimensional Object Transformations.

## Create file path

Create file path to save hex sticker to.

```{r}
filename <- here::here("inst","hex","hex.png")
dir.create(dirname(filename), showWarnings = FALSE, recursive = TRUE)
```


## Create cube hex

```{r}
#### Install dep ####
if(!require("threed")){
    remotes::install_github("coolbutuseless/threed")
}
#### Change view perspective ####
camera_to_world <- threed::look_at_matrix(eye = c(4, 4, 4), 
                                          at = c(0, 0, 0)) 
obj <- threed::mesh3dobj$cube |>
    threed::rotate_by(angle = 0, v = c(0, 1, 0)) |>
    threed::transform_by(threed::invert_matrix(camera_to_world)) |>
    threed::orthographic_projection()
cube_df <- as.data.frame(obj) |>
  dplyr::mutate(x=scales::rescale(x,c(-20,120)),
                y=scales::rescale(y,c(-20,120)))  

```

## Get coordinates

### Hexagon

```{r}
polygon_df <- dplyr::tibble(
  angle = seq(0, 2*pi, length.out = 7) + pi/6,
  x = cos(angle),
  y = sin(angle)
) |>
  dplyr::mutate(x=scales::rescale(x,c(-25,125)),
                y=scales::rescale(y,c(-25,125))) 
```


### UMAP

```{r}
obj1 <- TabulaMurisData::TabulaMurisSmartSeq2()
obj2 <- scKirby::se_to_seurat(obj1)
obj2 <- scKirby::process_seurat(obj2)
umap <- scKirby::get_obsm(obj2,keys = "umap")[[1]] |>
  data.frame(row.names=NULL) |>
              `colnames<-`(c("x","y")) |>
  dplyr::mutate_all(.funs = scales::rescale,c(0,100))

```

### Text 

```{r}
textPlot <- function(title,
                     subtitle=NULL,
                     plotname=tempfile(),
                     cex=22){
  tmp <- paste0(plotname, ".pdf")
  graphics::par(mar=c(0,0,0,0))
  grDevices::pdf(tmp,width = 12, height = 12)
  plot(c(0, 1), c(0, 1), ann = FALSE, bty = 'n', type = 'n', xaxt = 'n', yaxt = 'n')
  graphics::text(x = 0.5, y = 0.5, paste(title), 
                 cex = cex, col = "black",
                 family="serif", font=2, adj=0.5)
  if(!is.null(subtitle)){
      graphics::text(x = 0.5, y = 0.25, paste(subtitle), 
                     cex = cex/5, col = "black",
                     family="serif", font=2, adj=0.5) 
  }
  grDevices::dev.off()
  return(tmp)
}
tmp <- textPlot(plotname = here::here("inst/hex/scNLP_text"),
                title = "scNLP",
                subtitle="Natural Language Processing\nfor Single-cell Omics") 


img <- magick::image_read(tmp) |>
  magick::image_quantize(max=2, colorspace = 'gray', dither=TRUE) |>
  magick::image_scale(geometry = magick::geometry_size_pixels(width=300, 
                                                              height=300, 
                                                              preserve_aspect=TRUE)) |> magick::image_rotate(degrees = 90)
# Image manipulation
mat <- t(1L - 1L * (img[[1]][1,,] > 180))
img_df <- data.frame(which(mat>0,arr.ind = TRUE)) |>
  `colnames<-`(c("x","y")) |>
  dplyr::mutate(x = scales::rescale(x,c(-10,110)),
                y = scales::rescale(y,c(10,90))
                )

df <- rbind(
  cbind(id=seq_len(nrow(img_df)),
        umap[seq_len(nrow(img_df)),],
        state="text"),
  cbind(id=seq_len(nrow(img_df)),
        img_df,
        state="umap")
  )
```

## Plot

```{r}    
gp <- ggplot2::ggplot(df,aes(x,y, group=id, color=y)) +
  geom_polygon(data = polygon_df,
               aes(x,y),
               color="white", 
               fill=ggplot2::alpha("black",.85),
               inherit.aes = FALSE) +
  geom_point(show.legend = FALSE, size=.5) +
  scale_color_viridis_c(alpha = .9,option = "plasma", direction = -1) +
  ggplot2::geom_polygon(
      data = cube_df,
      ggplot2::aes(x = x, y = y, 
                   group = zorder, 
                   linetype = hidden,  size = hidden,
                   fill = 0.5 * fnx + fny), 
      color = ggplot2::alpha("white",.25),
      size = .25, inherit.aes = FALSE,
      # alpha=1,
      show.legend = FALSE) +
    ggplot2::scale_linetype_manual(values = c('TRUE' = "FF", 'FALSE' = 'solid')) +
    
    ggplot2::scale_fill_gradientn(
        colors = ggplot2::alpha(rev(RColorBrewer::brewer.pal(9, "Greys")), 
                                alpha = .04)) +  
    ggplot2::coord_equal() +  
  theme_void() +
  scale_x_continuous(limits = c(-25,125)) +
  scale_y_continuous(limits = c(-25,125)) +
  gganimate::transition_states(state,
                               transition_length = 3, 
                               state_length = c(2,5)) +
  gganimate::ease_aes('linear') 

gganimate::anim_save(filename = here::here("inst/hex/scNLP.gif"),
                     gp)
```


## Using `hexSticker::geom_hexagon`

```{r} 
gghex <- 
    hexSticker::ggplot() +
    hexSticker::geom_hexagon(fill = ggplot2::alpha("black",.75), 
                             color = ggplot2::alpha("white",.5)) +
    hexSticker::geom_pkgname(package = pkg, 
                             size=20, 
                             color = ggplot2::alpha("white",.5)) +
    hexSticker::theme_sticker()  

hexSticker::save_sticker(filename = filename,
                         sticker = gghex)
print(gghex)
```


# Session Info 

<details> 

```{r Session Info}
utils::sessionInfo()
```

</details>  
